Understanding models for spatial structure in small-area estimation

### 2023-05-31 src/docs_paper/paper.Rmd
Will Adam ever finish writing this paper? Stay tuned to find out (I hope so)

---
title: Understanding models for spatial structure in small-area estimation
classoption: ba
author:
  - firstname: Adam 
    lastname: Howes
    email: ath19@ic.ac.uk
    affiliationref: A,B
    footnoterefs: 
      - t1
  - firstname: Jeffrey W.
    lastname: Eaton
    affiliationref: C,B
  - firstname: Seth R.
    lastname: Flaxman
    affiliationref: D
affiliations:
  - ref: A
    name: Imperial College London 
    address: Department of Mathematics
  - ref: B
    name: Imperial College London
    address: MRC Centre for Global Infectious Disease Analysis
  - ref: C
    name: Harvard University
    address: Center for Communicable Disease Dynamics, Harvard T.H. Chan School of Public Health
  - ref: D
    name: University of Oxford
    address: Department of Computer Science 
footnotes:
  - ref: t1
    text: "Corresponding Author"
supplements:
  - title: Title of Supplement A
    description: Short description of Supplement A.
  - title: Title of Supplement B
    description: Short description of Supplement B.
acknowledgements: |
  | Yidan Xu undertook the Mary Lister McCammon Summer Research Fellowship 2019 on related work. AH thanks Robert Ashton for help with \textsf{R} package development, and Harrison Zhu, Theo Rashid, Samir Bhatt, Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for helpful conversations.
abstract: |
  |  Spatial correlation between small-areas is typically accounted for using spatial random effects. However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions for non-grid geometries. Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous geostatistical data. Defining spatial random effects which instead correspond to aggregated Gaussian processes gives a more intuitively convincing correlation structure, and allows tools from kernel methods to be used in models for areal data.
  |
  |  In a simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications. The simulations were performed under model misspecification and across a range of vignette and realistic geometries with different regularities. We also compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
date: "`r Sys.Date()`"
bibliography: citations.bib
#linenumbers: true
#numbersections: true
output: 
  bookdown::pdf_book: 
    base_format: rticles::isba_article
    md_extensions: -autolink_bare_uris #keep these off or \printead{} will fail on email addresses
    includes:
      in_header: preamble.tex
---

@@

Within the GRMF framework, attempts to more accurately reflect irregular spatial structure have defined weights specifying the extent to which neighbours are related.
@duncan2017spatial compared seventeen methods for specifying these weights and did not find any which outperform the Besag model, though this conclusion was specific to the often-analysed Scottish lip cancer example, and based on the deviance information criteria (DIC) which is recommended against by @vehtari2017practical.
Another approach is to take into account that areal data is typically produced by aggregating higher resolution data.
Obtaining inference about a variable at a different resolution to that it was observed at is known in geostatistics as the change-of-support problem [@gelfand2001change], and dates to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
Ideas from change-of-support modelling are applied by @kelsall2002modeling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
This type of model is particularly natural in the log-Gaussian Cox process modelling framework [@li2012log; @diggle2013spatial; @taylor2015bayesian; @johnson2019spatially].
<!-- `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages. -->
Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2015tmb], and made available as a part of the `disaggregation` package [@nandi2020disaggregation].

Recognising that spatial heterogeneity in outcomes commonly reflects a combination of both spatially correlated processes and specific circumstances at a given location, it is usually recommended to use a random effect specification which includes both spatially structured and unstructured components.
Examples include the BYM2 model [@simpson2017penalising] and earlier convolutions such as the BYM [@besag1991bayesian], Leroux [@leroux2000estimation] or Dean [@dean2001detecting] models.
Evidence for this recommendation includes the comparison studies of @lee2011comparison and @riebler2016intuitive.
Improvements to the Besag model are likely transferable to improving these convolution models.
The Naomi HIV small-area estimation model [@eaton2021naomi], which motivates our work, uses both Besag and BYM2 spatial random effects, at different levels of the hierarchical model, and both models are of substantive practical interest.

The remainder of this paper is organised as follows.
Section \ref{sec:background} provides background on areal data and the Bayesian hierarchical modelling approach to small-area estimation.
In Section \ref{sec:ab}, we review developments in specifying spatial random effects based on adjacency, before presenting an alternative approach based on kernels in Section \ref{sec:ks}.
In Section \ref{sec:simulation} we compare models using simulated data, before applying them to mapping HIV prevalence in sub-Saharan Africa in Section \ref{sec:hiv}.
Finally, we discuss our conclusions and directions for future research in Section \ref{sec:discussion}.

@@

# Spatial random effects using adjacency \label{sec:ab}

@@

Often, as well as spatial structure, there exists IID over-dispersion in the residuals and it is inappropriate to use purely spatially structured random effects in the model.
The Besag-York-Mollié (BYM) model of @besag1991bayesian, accounts for this in a natural way by decomposing the spatial random effect $\bphi = \bv + \bu$ into a sum of an unstructured IID component $\bv$ and a spatially structured Besag component $\bu$, each of which with their own respective precision parameters $\tau_v$ and $\tau_u$.
The resulting distribution is
\begin{equation}
    \bphi \sim \N(0, \tau_v^{-1} \mathbf{I} + \tau_u^{-1} \mathbf{R}^{-}) \label{eq:bym}.
\end{equation}
Including both $\bv$ and $\bu$ is intended to enable the model to learn the relative extent of the unstructured and structured components via $\tau_v$ and $\tau_u$.
However, in this specification, scaling of the Besag precision matrix $\mathbf{Q}$ is not taken into account despite this issue being particularly pertinent when dealing with multiple sources of noise.
In particular, placing a joint prior $(\tau_u, \tau_v) \sim p(\tau_u, \tau_v)$ which doesn't privilege either component is more easily accomplished if $\mathbf{Q}$ and $\mathbf{I}$ have the same scale.
Additionally, supposing we have a prior belief that the over-dispersion is primarily IID and $\bv$ accounts for the majority of the dispersion, then it is not immediately obvious how to represent this belief using $p(\tau_u, \tau_v)$, without inadvertently altering the prior about the overall variation.
This highlights identifiability issues of the parameters $(\tau_u, \tau_v)$ resulting from them not being orthogonal.
Building on the models of @leroux2000estimation and @dean2001detecting which tackle this the identifiability problem, but do not scale the spatially structured noise, @simpson2017penalising propose a reparameterisation $(\tau_v, \tau_u) \mapsto (\tau_\phi, \pi)$ of the BYM model known as the BYM2 model and given by
\begin{align}
\bphi = \frac{1}{\tau_\phi} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right), \label{eq:bym2}
\end{align}
where $\tau_\phi$ is the marginal precision of $\bphi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $\bu^\star$ is a scaled version of $u$ with precision matrix given by the scaled structure matrix $\mathbf{R}^\star$ (see Appendix).
When $\pi = 0$ the random effects are IID, and when $\pi = 1$ the random effects follow the Besag model.
To borrow an analogy, the parameterisation $(\tau_v, \tau_u)$ is like having one hot water and one cold water tap, whereas the parameterisation $(\tau_\phi, \pi)$ is like a mixer tap where the amount of water and its temperature can be adjusted separately [@rue2020comment].

# Spatial random effects using kernels \label{sec:ks}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{depends/integration-method.pdf}
\caption{Three ways to choose integration points with $L_i = 10$ for all $i$ in Malawi, as implemented by \texttt{sf::st\_sample}.}
\label{fig:41}
\end{figure}

@@

@@

If areal kernel spatial random effects are created using the integrated kernel, then aggregation occurs at the level of the latent field in the three stage model rather than at the level of the data.
This corresponds to a generative model of the form $y_i \sim p(y_i \, | \,  g^{-1}(x_i))$, with $x_i = |A_i|^{-1}\int_{A_i} x(s) \text{d} s$.
If the link function $g$ is the identity or linear then aggregation at the level of the latent field is equivalent to aggregation at the level of the data.
On the other hand, for non-linear link functions $g$ such as the exponential or logistic, the generative model does not match the proposed data generating process $y_i = |A_i|^{-1}\int_{A_i} y(s) \text{d}s$.
<!-- Add equations for mean $\mathbb{E}(y_i)$ and variance $\mathbb{V}(y_i)$ of under this model, as a function of $g$. -->
<!-- Explain under what circumstances we believe this approximation to be acceptable. -->

Most of the time we do not expect to be able to calculate the integral in Equation \ref{eq:ik2} analytically.
Instead, given $n$ collections of $L_i$ samples $\{s^{(i)}_l \}_{l = 1}^{L_i} \sim \mathcal{U}(A_i)$ drawn uniformly from each area then the integral may be approximated using Monte Carlo by the double sum
\begin{equation}
K(A_i, A_j) \approx \frac{1}{L_i L_j} \sum_{l = 1}^{L_i} \sum_{m = 1}^{L_j}
w_i \left( s^{(i)}_l \right) w_j \left( s^{(j)}_m \right)
k \left( s^{(i)}_l, s^{(j)}_m \right). \label{eq:mcintegratedkernel}
\end{equation}
Equivalently, samples drawn from $W_i$ may be used without weighting by $w_i(s)$.
Integration points may also be selected by some deterministic process, and used in Equation \ref{eq:mcintegratedkernel} to give a numerical integration estimate of the kernel.
Taking any of these approaches requires $\mathcal{O}(\sum_{i = 1}^n \sum_{j = 1}^n L_i L_j)$ evaluations of the kernel $k$ to compute the $n \times n$ Gram matrix $K$.
This imposes a significant computational cost if the Gram matrix is often recomputed during inference, as is the case in MCMC when any of the kernel hyperparameters are learnt, placing a limit on the number of samples or integration points it is feasible to use.
To make inference more feasible, @kelsall2002modeling reduce the number of Gram matrix constructions and inversions required, by using a discrete hyperparameter prior.

@@

Disaggregation regression, also known as downscaling or interpolation, is another closely related approach.
Rather than focusing on the aggregate nature of areal observations as primarily a route towards better area-level estimates, disaggregation regression aims to producing high-resolution or point-level estimates from areal observations [@utazi2019spatial; @nandi2020disaggregation].
<!-- Accurate point-level estimates are close to being a necessary intermediate goal towards obtaining accurate area-level estimates. -->
<!-- Disaggregation is challenging without auxiliary information. -->
<!-- A unified view on Bayesian varying coefficient models by Franco-Villoria et al. -->

@@

\begin{figure}
\centering
\includegraphics{depends/simulation-geometries.pdf}
\caption{The seven simulation geometries that we considered.}
\label{fig:51}
\end{figure}

We tested the ability of inferential models with varying spatial random effect specifications to accurately recover small-area quantities.
The data and modelling choices were designed with a spatial epidemiology setting in mind.
The \textsf{R} [@R2021] code used is available from `github.com/athowes/areal-comparison`.
We used the `orderly` package [@orderly] for reproducible research.

@@

\begin{table}[]
\small
\begin{tabularx}{\textwidth}{llX}
\toprule
 & Simulation model & Additional details \\
\midrule
\hypertarget{S1}{\texttt{S1}} & IID & $\phi \sim \mathcal{N}(0, I_n)$ \\
\hypertarget{S2}{\texttt{S2}} & Besag & $\phi \sim \mathcal{N}(0, (R^\star)^{-})$ as described in Section \ref{sec:besag} where $R^\star$ is the scaled structure matrix \\
\hypertarget{S3}{\texttt{S3}} & IK & $\phi \sim \mathcal{N}(0, K^\star)$ as described in Section \ref{sec:ik} where $K_{ij}$ is calculated via Equation \ref{eq:mcintegratedkernel} using the Mat\'{e}rn kernel with $\nu = 3/2$, $l = 2.5$ and $L_i = 100$ samples are drawn uniformly from each area.
Like the Besag model, we scaled the Gram matrix to have generalised variance one \\
\bottomrule
\end{tabularx}
\caption{The three simulation models from which we generate $\phi$.}
\label{tab:simulationmodels}
\small
\end{table}

To study robustness to spatial misspecification we simulated synthetic data-sets from three known data generating processes, using spatial random effects $\bphi$ generated according to IID, Besag and integrated kernel (IK) simulation models (Table \ref{tab:simulationmodels}).
We then generated synthetic data $\y = (y_1, \ldots, y_n)^\top$ of a form analogous to a household survey whereby $y_i \sim \text{Bin}(m_i, \rho_i)$ where the probabilities $\rho_i \in [0, 1]$ are linked to the latent random variables $x_i \in \mathbb{R}$ via
\begin{equation}
\log \left( \frac{\rho_i}{1 - \rho_i} \right) = x_i = \beta_0 + \phi_i, \quad i = 1, \ldots, n.
\end{equation}
The sample sizes were fixed as $m_i = 25$ for all $i$, the intercept parameter as $\beta_0 = -2$ and the spatial random effect precision parameter as $\tau_\phi = 1$.
We used seven different geometries: the four vignette geometries sharing an adjacency graph in Figure \ref{fig:tikz1}, as well as three more realistic geometries, a $6 \times 6$ lattice grid, the 33 districts of Côte d'Ivoire and the 36 congressional districts of Texas (Figure \ref{fig:51}).
The more realistic geometries were chosen to represent variation over spatial regularity, allowing us to test how model performance varies by geometry regularity.
For each combination of spatial random effect and geometry we replicated the simulation process above 200 times to generate a total of 1400 synthetic data-sets.

@@

\begin{table}
\small
\begin{tabularx}{\textwidth}{lllX}
\toprule
& Inferential model & Algorithm & Additional details \\
 \midrule
\hypertarget{I1}{\texttt{I1}} & Constant & AGHQ &
No spatial random effects \label{row:I1} \\
\hypertarget{I2}{\texttt{I2}} & IID & AGHQ &
$\phi \sim \mathcal{N}(0, \tau_\phi^{-1} I_n)$\\
\hypertarget{I3}{\texttt{I3}} & Besag & AGHQ &
$\phi \sim \mathcal{N}(0, (\tau_\phi R^\star)^{-})$ following additional recommendations of \citet{freni2018note}  \\
\hypertarget{I4}{\texttt{I4}} & BYM2 & AGHQ &
$\phi = \tau_\phi^{-1} \left( \sqrt{1- \pi} \, v + \sqrt{\pi} \, u^\star \right)$ with a $\text{Beta}(0.5, 0.5)$ prior on $\pi \in [0, 1]$ \\
\hypertarget{I5}{\texttt{I5}} & FCK & AGHQ &
$\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = k(c_i, c_j)$ where the length-scale $l$ is fixed \\
\hypertarget{I6}{\texttt{I6}} & CK & AGHQ & As in model \hypertarget{I5}{I5}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
\hypertarget{I7}{\texttt{I7}} & FIK & AGHQ & $\phi \sim \mathcal{N}(0, \tau_\phi^{-1} K)$ with $K_{ij} = \frac{1}{10^2} \sum_{l = 1}^{10} \sum_{m = 1}^{10} k \left( s^{(i)}_l, s^{(j)}_m \right)$ \newline with hexagonal integration point spacing and fixed length-scale $l$ \\
\hypertarget{I8}{\texttt{I8}} & IK & AGHQ & As in model \hypertarget{I7}{I7}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
\bottomrule
\end{tabularx}
\caption{All inferential models are implemented as a part of the \textsf{R} package \texttt{arealutils} available from \url{github.com/athowes/arealutils}.}
\label{tab:inferentialmodels}
\small
\end{table}

@@

We placed a half-Gaussian prior on the standard deviation [@gelman2006prior] such that $\sigma_\phi \sim \mathcal{N}_+(0, 2.5^2)$.
The weakly informative value 2.5 was chosen to avoid placing significant prior density on the region $\sigma_\phi > 5$, which after a logistic transformation facilitates variation on the probability scale very close to either zero or one, neither of which we assumed to be plausible.
A weakly informative $\mathcal{N}(-2, 1)$ prior was placed on $\beta_0$, setting most of the prior probability density for $\rho_i$ within a range $[0, 0.25]$ typical for a disease prevalence.

@@

We used adaptive Gauss-Hermite quadrature (AGHQ) with $k = 3$ quadrature points per hyperparamter dimension to perform approximate Bayesian inference.

@@

We assessed the coverage of our estimates using posterior predictive checks based on the quantile $q_i = F(\omega_i)$ of the true value within each marginal posterior predictive distribution, computed using numerical integration.
For calibrated models, over repeated simulations, $q_i \sim \mathcal{U}[0, 1]$ such that at any given nominal coverage $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$ is also be $1 - \alpha$.
To check uniformity we used probability integral transform (PIT) histograms [@dawid1984present] and empirical cumulative distribution function (ECDF) difference plots [@aldor2013power; @sailynoja2021graphical].

>>>>

---
title: Understanding models for spatial structure in small-area estimation
classoption: ba
author:
  - firstname: Adam 
    lastname: Howes
    email: ath19@ic.ac.uk
    affiliationref: A,B
    footnoterefs: 
      - t1
  - firstname: Jeffrey W.
    lastname: Eaton
    affiliationref: C,B
  - firstname: Seth R.
    lastname: Flaxman
    affiliationref: D
affiliations:
  - ref: A
    name: Imperial College London 
    address: Department of Mathematics
  - ref: B
    name: Imperial College London
    address: MRC Centre for Global Infectious Disease Analysis
  - ref: C
    name: Harvard University
    address: Center for Communicable Disease Dynamics, Harvard T.H. Chan School of Public Health
  - ref: D
    name: University of Oxford
    address: Department of Computer Science 
footnotes:
  - ref: t1
    text: "Corresponding Author"
supplements:
  - title: Appendix
    description: All further materials for this study.
acknowledgements: |
  | AH thanks Robert Ashton for help with \textsf{R} package development, and Harrison Zhu, Theo Rashid, Tim Wolock, Samir Bhatt, Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for helpful conversations. Yidan Xu undertook the Mary Lister McCammon Summer Research Fellowship 2019 on related work.
abstract: |
  |  Spatial correlation between small-areas is typically accounted for using spatial random effects. However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions for non-grid geometries. Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous geostatistical data. Defining spatial random effects which instead correspond to aggregated Gaussian processes gives a more intuitively convincing correlation structure, and allows tools from kernel methods to be used in models for areal data.
  |
  |  In a simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications. The simulations were performed under model misspecification and across a range of vignette and realistic geometries with different regularities. We also compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
date: "`r Sys.Date()`"
bibliography: citations.bib
#linenumbers: true
#numbersections: true
output: 
  bookdown::pdf_book: 
    base_format: rticles::isba_article
    md_extensions: -autolink_bare_uris #keep these off or \printead{} will fail on email addresses
    includes:
      in_header: preamble.tex
---

@@

Within the GRMF framework, attempts to more accurately reflect irregular spatial structure have defined weights specifying the extent to which neighbours are related.
@duncan2017spatial compared seventeen methods for specifying these weights, but did not find any which outperform the Besag model.
This conclusion was specific to the often-analysed Scottish lip cancer example, and based on the deviance information criteria (DIC) which is recommended against by @vehtari2017practical.
Another approach is to take into account that areal data is typically produced by aggregating higher resolution data.
Obtaining inference about a variable at a different resolution to that it was observed at is known in geostatistics as the change-of-support problem [@gelfand2001change], and dates to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
@kelsall2002modeling applied ideas from change-of-support modelling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
This type of model is particularly natural in the log-Gaussian Cox process modelling framework [@li2012log; @diggle2013spatial; @taylor2015bayesian; @johnson2019spatially].
<!-- `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages. -->
Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2016tmb], and made available as a part of the `disaggregation` package [@nandi2020disaggregation].

Spatial heterogeneity commonly reflects a combination of both spatially correlated processes and specific circumstances at a given area.
It is usually therefore recommended to use a spatial random effect specification which includes both structured and unstructured components, such as the BYM2 model [@simpson2017penalising] or earlier convolutions such as the BYM [@besag1991bayesian], Leroux [@leroux2000estimation] or Dean [@dean2001detecting] models.
Evidence for this recommendation includes the comparison studies of @lee2011comparison and @riebler2016intuitive.
Any improvement to the Besag model is likely directly transferable to improving these convolution models, and both are of substantive practical interest.
For example, the Naomi HIV small-area estimation model [@eaton2021naomi] uses both Besag and BYM2 random effects.

The remainder of this paper is organised as follows.
Section \ref{sec:background} provides background on areal data and the Bayesian hierarchical modelling approach to small-area estimation.
In Section \ref{sec:ab}, we review developments in specifying spatial random effects based on adjacency, before presenting an alternative approach based on kernels in Section \ref{sec:ks}.
In Section \ref{sec:simulation} we present a simulated case-study, before moving on to mapping HIV prevalence in sub-Saharan Africa in Section \ref{sec:hiv}.
Finally, we discuss our conclusions and directions for future research in Section \ref{sec:discussion}.

@@

# Spatial random effects defined using adjacency \label{sec:ab}

@@

Often, as well as spatial structure, there exists IID over-dispersion in the residuals and it is inappropriate to use purely spatially structured random effects in the model.
The Besag-York-Mollié (BYM) model of @besag1991bayesian, accounts for this in a natural way by decomposing the spatial random effect $\bphi = \bv + \bu$ into a sum of an unstructured IID component $\bv$ and a spatially structured Besag component $\bu$, each of which with their own respective precision parameters $\tau_v$ and $\tau_u$.
The resulting distribution is
\begin{equation}
    \bphi \sim \N(0, \tau_v^{-1} \mathbf{I} + \tau_u^{-1} \mathbf{R}^{-}) \label{eq:bym}.
\end{equation}
Including both $\bv$ and $\bu$ is intended to enable the model to learn the relative extent of the unstructured and structured components via $\tau_v$ and $\tau_u$.
However, in this specification, scaling of the Besag precision matrix $\mathbf{Q}$ is not taken into account despite this issue being particularly pertinent when dealing with multiple sources of noise.
In particular, placing a joint prior $(\tau_u, \tau_v) \sim p(\tau_u, \tau_v)$ which doesn't privilege either component is more easily accomplished if $\mathbf{Q}$ and $\mathbf{I}$ have the same scale.
Additionally, supposing we have a prior belief that the over-dispersion is primarily IID and $\bv$ accounts for the majority of the dispersion, then it is not immediately obvious how to represent this belief using $p(\tau_u, \tau_v)$, without inadvertently altering the prior about the overall variation.
This highlights identifiability issues of the parameters $(\tau_u, \tau_v)$ resulting from them not being orthogonal.
Building on the models of @leroux2000estimation and @dean2001detecting which tackle this the identifiability problem, but do not scale the spatially structured noise, @simpson2017penalising propose a reparameterisation $(\tau_v, \tau_u) \mapsto (\tau_\phi, \pi)$ of the BYM model known as the BYM2 model and given by
\begin{align}
\bphi = \frac{1}{\tau_\phi} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right), \label{eq:bym2}
\end{align}
where $\tau_\phi$ is the marginal precision of $\bphi$, $\pi \in [0, 1]$ gives the proportion of the marginal variance explained by each component, and $\bu^\star$ is a scaled version of $u$ with precision matrix given by the scaled structure matrix $\mathbf{R}^\star$ (see Appendix).
When $\pi = 0$ the random effects are IID, and when $\pi = 1$ the random effects follow the Besag model.
To borrow an analogy [@rue2020comment] the parameterisation $(\tau_v, \tau_u)$ is like having one hot water and one cold water tap, whereas the parameterisation $(\tau_\phi, \pi)$ is like a mixer tap where the amount of water and its temperature can be adjusted separately.

# Spatial random effects defined using kernels \label{sec:ks}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{depends/integration-method.pdf}
\caption{Three ways to choose nodes with $L_i = 10$ for all $i$ in Malawi, as implemented by \texttt{sf::st\_sample}.}
\label{fig:41}
\end{figure}

@@

### Accounting for heterogeneity

@@

### Computation

Most of the time we do not expect to be able to calculate Equation \ref{eq:ik2} analytically.
Instead, given $n$ collections of $L_i$ samples $\{s^{(i)}_l \}_{l = 1}^{L_i} \sim \mathcal{U}(A_i)$ drawn uniformly from each area then the integral may be approximated using Monte Carlo by the double sum
\begin{equation}
K(A_i, A_j) \approx \frac{1}{L_i L_j} \sum_{l = 1}^{L_i} \sum_{m = 1}^{L_j}
w_i \left( s^{(i)}_l \right) w_j \left( s^{(j)}_m \right)
k \left( s^{(i)}_l, s^{(j)}_m \right). \label{eq:mcintegratedkernel}
\end{equation}
Equivalently, samples drawn from $W_i$ may be used without weighting by $w_i(s)$.
Nodes may also be selected by some deterministic process to give a numerical quadrature estimate of the kernel.
These approaches require $\mathcal{O}(\sum_{i = 1}^n \sum_{j = 1}^n L_i L_j)$ evaluations of the kernel $k$ to compute the $n \times n$ Gram matrix $K$.
This imposes a significant computational cost if the Gram matrix is often recomputed during inference, as is the case in MCMC when any of the kernel hyperparameters are learnt, placing a limit on the number of samples or nodes it is feasible to use.
@kelsall2002modeling make inference more feasible by using a discrete hyperparameter prior to reduce the number of Gram matrix constructions and inversions required.

### Mismatch to data generating process

Aggregation via the integrated kernel occurs at the level of the latent field rather than at the level of the data.
<!-- This corresponds to a generative model of the form $y_i \sim p(y_i \, | \,  g^{-1}(x_i))$, with $x_i = |A_i|^{-1}\int_{A_i} x(s) \text{d} s$. -->
If the link function $g$ is the identity or linear then aggregation at the level of the latent field is equivalent to aggregation at the level of the data.
On the other hand, for non-linear link functions $g$ such as the exponential or logistic, the generative model does not match the proposed data generating process.
<!-- Add equations for mean $\mathbb{E}(y_i)$ and variance $\mathbb{V}(y_i)$ of under this model, as a function of $g$. -->
<!-- Explain under what circumstances we believe this approximation to be acceptable. -->

@@

Disaggregation regression, also known as downscaling or interpolation, is another closely related approach.
Rather than focusing on the aggregate nature of areal observations as primarily a route towards better area-level estimates, disaggregation regression aims to produce high-resolution or point-level estimates from areal observations [@utazi2019spatial; @nandi2020disaggregation].
<!-- Accurate point-level estimates are close to being a necessary intermediate goal towards obtaining accurate area-level estimates. -->
<!-- Disaggregation is challenging without auxiliary information. -->
<!-- A unified view on Bayesian varying coefficient models by Franco-Villoria et al. -->

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{depends/simulation-geometries.pdf}
\caption{The seven simulation geometries that we considered.}
\label{fig:51}
\end{figure}

We tested the ability of inferential models with varying spatial random effect specifications to accurately recover small-area quantities.
The data and modelling choices were designed with a spatial epidemiology setting in mind.
The \textsf{R} [@R2021] code used is available from `github.com/athowes/areal-comparison`.
We used `orderly` [@orderly] for reproducible research, `ggplot2` for data visualisation [@wickham2016ggplot2] and `rticles` [@allaire2022rticles] for reporting via `rmarkdown` [@allaire2022rmarkdown].

@@

\begin{table}[]
\small
\begin{tabularx}{\textwidth}{llX}
\toprule
 & Simulation model & Additional details \\
\midrule
\hypertarget{S1}{\texttt{S1}} & IID & $\bphi \sim \mathcal{N}(0, \mathbf{I}_n)$ \\
\hypertarget{S2}{\texttt{S2}} & Besag & $\bphi \sim \mathcal{N}(0, (\mathbf{R}^\star)^{-})$ as described in Section \ref{sec:besag} where $\mathbf{R}^\star$ is the scaled structure matrix \\
\hypertarget{S3}{\texttt{S3}} & IK & $\bphi \sim \mathcal{N}(0, \mathbf{K}^\star)$ as described in Section \ref{sec:ik} where $K_{ij}$ is calculated via Equation \ref{eq:mcintegratedkernel} using the Mat\'{e}rn kernel with $\nu = 3/2$, $l = 2.5$ and $L_i = 100$ samples are drawn uniformly from each area.
Like the Besag model, we scaled the Gram matrix to have generalised variance one \\
\bottomrule
\end{tabularx}
\caption{The three simulation models from which we generate $\bphi$.}
\label{tab:simulationmodels}
\small
\end{table}

To study robustness to spatial misspecification we simulated synthetic data-sets from three known data generating processes, using spatial random effects $\bphi$ generated according to IID, Besag and integrated kernel (IK) simulation models (Table \ref{tab:simulationmodels}).
We then generated synthetic data $\y = (y_i)_{i \in [n]}$ of a form analogous to a household survey whereby $y_i \sim \text{Bin}(m_i, \rho_i)$ where the probabilities $\rho_i \in [0, 1]$ are linked to the linear predictors $\eta_i \in \mathbb{R}$ via
\begin{equation}
\log \left( \frac{\rho_i}{1 - \rho_i} \right) = \eta_i = \beta_0 + \phi_i, \quad i \in [n].
\end{equation}
The sample sizes were fixed as $m_i = 25$ for all $i \in [n]$, the intercept parameter as $\beta_0 = -2$ and the spatial random effect precision parameter as $\tau_\phi = 1$.
We used seven different geometries including the four vignette geometries sharing an adjacency graph (Figure \ref{fig:tikz1}), as well as three more realistic geometries, a $6 \times 6$ lattice grid, the 33 districts of Côte d'Ivoire and the 36 congressional districts of Texas (Figure \ref{fig:51}).
The more realistic geometries were chosen to represent variation over spatial regularity, allowing us to test how model performance varies by geometry regularity.
For each combination of spatial random effect and geometry we replicated the simulation process above 200 times to generate a total of 1400 synthetic data-sets.

@@

\begin{table}
\small
\begin{tabularx}{\textwidth}{llX}
\toprule
& Inferential model & Additional details \\
 \midrule
\hypertarget{I1}{\texttt{I1}} & Constant &
No spatial random effects \label{row:I1} \\
\hypertarget{I2}{\texttt{I2}} & IID &
$\bphi \sim \mathcal{N}(0, \tau_\phi^{-1} \mathbf{I}_n)$\\
\hypertarget{I3}{\texttt{I3}} & Besag &
$\bphi \sim \mathcal{N}(0, (\tau_\phi \mathbf{R}^\star)^{-})$ following additional recommendations of \citet{freni2018note}  \\
\hypertarget{I4}{\texttt{I4}} & BYM2 &
$\bphi = \tau_\phi^{-1} \left( \sqrt{1- \pi} \, \bv + \sqrt{\pi} \, \bu^\star \right)$ with a $\text{Beta}(0.5, 0.5)$ prior on $\pi \in [0, 1]$ \\
\hypertarget{I5}{\texttt{I5}} & FCK &
$\bphi \sim \mathcal{N}(0, \tau_\phi^{-1} \mathbf{K})$ with $K_{ij} = k(c_i, c_j)$ where the length-scale $l$ is fixed \\
\hypertarget{I6}{\texttt{I6}} & CK & As \hypertarget{I5}{\texttt{I5}}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
\hypertarget{I7}{\texttt{I7}} & FIK & $\bphi \sim \mathcal{N}(0, \tau_\phi^{-1} \mathbf{K})$ with $K_{ij} = \frac{1}{10^2} \sum_{l = 1}^{10} \sum_{m = 1}^{10} k \left( s^{(i)}_l, s^{(j)}_m \right)$ \newline with hexagonal integration point spacing and fixed length-scale $l$ \\
\hypertarget{I8}{\texttt{I8}} & IK & As \hypertarget{I7}{\texttt{I7}}, with length-scale prior $l \sim \mathrm{Inv{\text-}Gamma}(a, b)$ \\
\bottomrule
\end{tabularx}
\caption{All inferential models are implemented as a part of the \textsf{R} package \texttt{arealutils} available from \url{github.com/athowes/arealutils}. The AGHQ algorithm was used for approximate Bayesian inference of all inferential models.}
\label{tab:inferentialmodels}
\small
\end{table}

@@

We placed a half-Gaussian prior on the standard deviation [@gelman2006prior] such that $\sigma_\phi \sim \mathcal{N}_+(0, 2.5^2)$.
The weakly informative value 2.5 was chosen to avoid placing significant prior density on the region $\sigma_\phi > 5$, which after a logistic transformation facilitates variation on the probability scale very close to either zero or one.
A weakly informative $\mathcal{N}(-2, 1)$ prior was placed on $\beta_0$, setting most of the prior probability density for $\rho_i$ within a range $[0, 0.25]$ typical for a disease prevalence.

@@

We used adaptive Gauss-Hermite quadrature [AGHQ; @stringer2022fast] with $k = 3$ quadrature points per hyperparameter dimension to perform approximate Bayesian inference.
Each model is implemented using a Template Model Builder \textsc{C++} template for the log-posterior via the `TMB` package [@kristensen2016tmb]
We are then able to conduct inference using both AGHQ via the `aghq` package [@stringer2021implementing] and the No-U-Turn Sampling (NUTS) Hamiltonian Monte Carlo (HMC) algorithm using Stan [@carpenter2017stan] via the `tmbstan` package [@monnahan2018no].
In Appendix we demonstrate that the inference results from AGHQ are comparable to those obtained running NUTS over a longer time.
We also compare to empirical Bayes, and where possible integrated nested Laplace approximation via `R-INLA` [@martins2013bayesian].

@@

For calibrated models, over repeated simulations, $F(\omega_i) = q_i \sim \mathcal{U}[0, 1]$ such that at any given nominal coverage $1 - \alpha$, the proportion of quantile-based credible intervals containing the true value $\omega_i$ is also $1 - \alpha$.
We checked uniformity using probability integral transform (PIT) histograms [@dawid1984present] and empirical cumulative distribution function (ECDF) difference plots [@aldor2013power; @sailynoja2022graphical].
